---
title: "WUP_04"
author: "Leonel D. Bautista"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file="GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Using ChatGPT and CandP helper files to compare numerical data and categorical data in the GSS data set. We will compare WORDSUM and how it is effected by Education level.   

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r include=TRUE,echo=FALSE}
ggplot(data = mydata, aes(x = DEGREE, y = WORDSUM) ) +
geom_boxplot ( notch=FALSE , outlier.colour = "red" , outlier.size = 2 , fill = "darkgray" ) +
stat_boxplot ( geom = 'errorbar' ) +
labs ( x = "DEGREE" , y = "WORDSUM" , title = "WORDSUM explained by DEGREE" ) +
theme ( plot.title = element_text (hjust = 0.5, size = 16) )
```   

It appears that the entire IQR of the High School and LT High School is underneath the median of all of the college groups.   

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r include=TRUE,echo=FALSE}
favstats( WORDSUM ~ DEGREE , data= mydata )
```

The mean and median of the LT High School and the High School subcategories are far lower than the means and medians of the univeristy level data points.   

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r include=TRUE,echo=FALSE}
model1 <- aov( WORDSUM ~ DEGREE , data = mydata )
summary(model1)
```
```{r include=TRUE,echo=FALSE}
par(mar = c(3, 8.7, 2, 2))
plot(TukeyHSD(model1, conf.level = 0.98),las=2)
```

The inferential analysis, which included an ANOVA test, provided critical insights into the relationship between education level and WORDSUM scores. The results of the ANOVA test yielded a significant F-value 0.000236, indicating that there is a statistically significant difference in mean WORDSUM scores across at least two levels of education.

Subsequently, the post-hoc Tukey's test was conducted to pinpoint specific differences between education levels. This additional analysis allowed us to identify which specific pairs of education levels exhibited significant disparities in mean WORDSUM scores.   


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variables causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

In conclusion, our findings strongly support the alternative hypothesis that there is a significant difference in average WORDSUM scores between at least two levels of education. This implies that education level is a factor influencing performance on the WORDSUM test.     

These results have broader implications, suggesting that individuals with higher educational attainment may, on average, perform better on this particular test. However, it is important to acknowledge that while we have identified a statistical association, causation cannot be determined from this analysis alone. Other unmeasured variables may contribute to the observed differences.   

Future research could delve into exploring the underlying mechanisms driving this relationship and consider additional factors that might mediate or moderate the impact of education on cognitive performance.   

Overall, this study contributes to our understanding of the interplay between education level and cognitive abilities, highlighting the potential significance of educational attainment in cognitive testing outcomes.   
